A New Hybrid Machine Learning Method for Stellar Parameter Inference

### 2025-01-26 paper/paper2/paper.tex
Up to 40/81 revisions

\subsection{Line Identification with \texttt{blase}}
The next step was to convert the PHOENIX subset into a physically interpretable intermediate representation: a table of spectral line properties rather than a\textbf{n} array of fluxes.
We used \texttt{blase}, which models spectral lines as Voigt profiles and tunes the profiles to mimic the original PHOENIX spectrum with back propagation.
Back propagation, put simply, is the process of moving in the autodiff gradient field to update ML model parameters (like mentioned earlier, usually against the gradient because we minimize loss functions).
Four parameters were optimized: the line center $\mu$, the log-amplitude $\ln(a)$, the Gaussian width $\sigma$, and the Lorentzian width $\gamma$.
The optimization used the Adam optimizer \textbf{\citep{adam}} with a learning rate of 0.05 over 100 epochs.
In addition, we limited two custom parameters: wing cut to 6000 and prominence to 0.005.
Wing cut \textbf{(in pixels)} is a parameter that determines the extent of the Voigt profile to evaluate, saving computational resources by not evaluating negligible line wings.
Prominence \textbf{(in normalized flux counts)} sets a lower limit for the amplitude of detected lines, which saves resources by disregarding shallow lines, so in our case we disregard lines that affect the spectrum by $<$ 0.5\%.
In short, our choices for wing cut and prominence decrease the computational cost of \texttt{blase}'s cloning process at the expense of decreasing its accuracy slightly.
We note that \texttt{blase} uses the pseudo-Voigt approximation, which saves on computational cost while remaining accurate to about 1\% \citep{pseudovoigt}.
The pseudo-Voigt approximation uses a weighted average of a Gaussian and Lorentzian as opposed to a convolution.
\texttt{blase}'s pseudo-Voigt profile implementation uses the following:
\begin{linenomath*}
\begin{subequations}
\begin{gather}
    \tilde{\mathsf{V}}_\mu(\lambda) = a\left[\eta \mathbf{L}(\lambda - \mu'; f) + (1 - \eta)\mathbf{G}(\lambda - \mu'; f)\right]\\
    \eta = \sum_{n=1}^3 \mathbf{u}_n\left(\dfrac{2\gamma}{f}\right)^n\\
    f = 32\sum_{n=0}^5 \mathbf{v}_n \left(\sqrt{2\ln(2)}\sigma\right)^{5-n}(\gamma)^n\\
    \nonumber \mathbf{u} = \mqty[1.36603 \\ -0.47719 \\ 0.11116] \quad \mathbf{v} = \mqty[1 \\ 2.69269 \\ 2.42843 \\ 4.47163 \\ 0.07842 \\ 1]
    \label{eq:equation2}
\end{gather}
\end{subequations}
\end{linenomath*}
where $\mathbf{L}$ and $\mathbf{G}$ are abbreviations for Lorentzian and Gaussian profiles, respectively.
Notice that we use $\mu'$ instead of $\mu$ in the formula.
This is because \texttt{blase} optionally allows the line center to shift slightly during optimization, and it is this shifted center which is used in computation.
The individual Voigt profiles are still indexed by $\mu$ for cross-model line identification, explained in the next section.
Once optimization was complete, the list of identified lines, was saved to a `state dictionary': a common representation for pre-trained machine learning models that can be stored to disk for reuse later.
These are stored in the \texttt{.pt} file format for each of the 1,314 PHOENIX subset grid points.
The total disk space these files take up is 465 MB (382 MB when downloaded as a zip archive).
For reference, the storage space the PHOENIX subset takes up on disk is approximately 8.1 GB.
This represents a data compression factor of around 20 just by recasting the spectrum with \texttt{blase}.

@@

Spectral lines were often only detected in some spectra from the PHOENIX subset.
In \autoref{fig:figure2}, we show that different grid points sport differing counts of detected spectral lines.
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure2}
    \caption{Number of detected spectral lines at each grid point of a slice of the PHOENIX subset at solar metallicity.
    We can see that the number of detected lines decreases with increasing $T_{\mathrm{eff}}$.
    Also note that from $T_{\mathrm{eff}} = 7000$ K onward, PHOENIX's sampling increment changes from 100 K to 200 K.}
    \label{fig:figure2}
\end{figure*}

This missing-lines phenomenon can arise for different reasons.
Perhaps for astrophysical reasons: stellar atmospheres genuinely do not produce that line at detectable strength at the given temperature and surface gravity.
Or alternatively, our line-finding and line-association algorithms missed it.
Whatever the cause, these missing lines have immediate practical consequences.
Rectilinear interpolation schemes break in regions where a line does not appear; you can't interpolate a quantity that simply doesn't exist.

To solve this, we artificially populated missing grid points with log-amplitudes of -1000, which retained interpolator stability while nullifying the evaluated line.
Examples of the appearance of missing sections in heatmaps where a line does not appear are shown in \autoref{fig:figure3} and \autoref{fig:figure4}.
In total, across the entire PHOENIX subset, \texttt{blase} detected 128,723 individual spectral lines.
Every one of these lines can be visualized as a manifold mapping a 3D stellar parameter vector to a 4D spectral line parameter vector, and every one must be interpolated.
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure3}
    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset slice at solar metallicity of an unknown spectral line.
    Notice the missing chunk in the top left of the figure; \texttt{blase} did not detect a spectral line here, but we have to artificially populate those points with lines that have $\ln(a) = -1000$.
    This and all spectral lines shown in this paper were identified using the NIST spectral line database \citep{NIST}.}
    \label{fig:figure3}
\end{figure*}
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure4}
    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset slice at solar metallicity of a Helium II spectral line.
    We can see that \texttt{blase} detects this line at only a select chunk of grid points in PHOENIX, leading to the large amount of missing data for the line.}
    \label{fig:figure4}
\end{figure*}

\subsection{Continuously Evaluable Manifolds}
For each line, the inputs to the interpolator were the three input parameters $T_{\mathrm{eff}}$, $\log(g)$, and [Fe/H], and the output was a list of four parameters, $\mu'$, $\ln(a)$, $\sigma$, and $\gamma$.
For each line, one of these interpolator objects was created using linear interpolation, and these interpolators were aggregated into a single list, which was then written to disk in the \texttt{.pkl} file storage format.
These interpolators generate multiple manifolds representing the following mapping:
\begin{gather}
    \mqty[T_{\mathrm{eff}} \\ \log(g) \\ \mathrm{[Fe/H]}] \rightarrow \mqty[\mu' \\ \ln(a) \\ \sigma \\ \gamma]
\end{gather}
These interpolators could now be evaluated at any point lying within the domain of the PHOENIX subset, turning a discretely sampled PHOENIX subset into a continuously evaluable function, sometimes called a spectral emulator.
With the given size of the PHOENIX subset, the interpolator list takes up 13.2 GB of disk space.
This evaluation is able to reconstruct an existing PHOENIX spectrum or alternatively interpolate a new spectrum within the domain of the PHOENIX subset, so we call this the PHOENIX generator.
In \autoref{fig:figure5} and \autoref{fig:figure6}, we show the same spectral lines as in \autoref{fig:figure3} and \autoref{fig:figure4}, but now supersampled using the PHOENIX generator evaluated over the same slice.
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure5}
    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ of our unknown spectral line vary over the PHOENIX subset slice at solar metallicity, now supersampled with the PHOENIX generator.
    Notice that the missing chunk in the top left still exists and does not display any artifacts, as the artificially populated points are removed after interpolation to retain the model's integrity.
    Also see that the x-axis spacing is now uniform, as the PHOENIX generator was evaluated at constant step.}
    \label{fig:figure5}
\end{figure*}
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure6}
    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ of our Helium II spectral line vary over the PHOENIX subset slice at solar metallicity, now supersampled with the PHOENIX generator.}
    \label{fig:figure6}
\end{figure*}

@@

\section{Bayesian Inference and Testing}
\label{sec:section4}
\subsection{Spectral Reconstruction Time}
A typical use case for the PHOENIX generator may be to batch reconstruct spectra from an array of input stellar parameters.
Therefore, there is some motivation to reduce the computational cost of this procedure to tractable levels.
We evaluate the computational time needed to use the PHOENIX generator in two distinct ways.
First, for a single input, which would be relevant in serial applications.
Second, for an array of multiple inputs, which would be relevant in parallel applications.
The \texttt{RegularGridInterpolator} API allows for the passing in of an entire array of input coordinates to be evaluated at once.
However using \texttt{blase} to reconstruct the spectrum from our interpolated state dictionary is always done serially, leading to what is actually more of a pseudo-parallel evaluation, but extremely performant nonetheless.
Performance results are shown in \autoref{fig:figure8}, and we can see that the multi-reconstruction is much faster than a series of single reconstructions.

\begin{figure}[t!]
    \centering
    \includegraphics[width=\columnwidth]{figure8}
    \caption{Line chart showing the time taken to reconstruct varying numbers of spectra using the PHOENIX generator (lower is better).
    We can see that the time taken per spectra for the serial implementation hovers around 15 seconds within run-to-run variance, while the pseudo-parallel implementation continually decreases in time taken per spectrum as the number of inputs increases.
    The speedup factor (higher is better) increases as more spectra are generated, which is also a desirable outcome.}
    \label{fig:figure8}
\end{figure}

\subsection{Inference Algorithm}
We elected to use Bayesian optimization as the inference algorithm, specifically the \texttt{gp\_minimize} function from the \texttt{scikit-optimize} library \citep{skopt}.
This algorithm uses a Gaussian Process to model the objective function, which in this case was the RMS (Root-Mean-Square) loss between the interpolated spectrum $\mathsf{M}$ and the true spectrum $\mathsf{D}$, defined as:
\begin{gather}
    \mathcal{L} = \ev{(\mathsf{M} - \mathsf{D})^{\circ 2}}^{1/2}
\end{gather}
The optimizer was configured to first run 100 random evaluations to seed the surrogate model, then run 20 more evaluations now guided by the surrogate model.
This totals to 120 evaluations, a large sample to create a fairly detailed surrogate model, then a moderately precise guided evaluation phase, which was deemed sufficient for this study.
Fine-tuning these numbers is possible, but simply not warranted for a proof-of concept method.
One inference run takes on average just under 7.5 minutes to complete.

@@

We know that in typical observational settings, a coarse estimate for $T_{\mathrm{eff}}$ tends to be fairly well-constrained from ancillary data prior to even looking at the spectra.
So when testing the inference algorithm, we limited its search space to only include $T_{\mathrm{eff}}$ which lay within 500 K of the true value on either side.
$\log(g)$ and [Fe/H] were allowed to vary freely.
The test sample consisted of 9 unique $T_{\mathrm{eff}}$ values, 3 unique $\log(g)$ values, and 2 unique [Fe/H] values, totaling 54 unique spectra in the test set.
$T_{\mathrm{eff}}$ ranged from 3000 K to 11000 K in increments of 1000 K, $\log(g)$ ranged from 2 to 6 in increments of 2, and [Fe/H] ranged from -0.5 to 0 in increments of 0.5.

@@

\subsection{Technical Considerations}
To reiterate, the manifold fitting steps are not end-to-end autodifferentiable.
As seen in \autoref{fig:figure1}, these steps rely on \texttt{scipy}, which is not equipped with autodiff.
Without autodifferentiability, the ability to ``machine learn'' is significantly reduced compared to a hypothetical monolithic JAX or PyTorch system.
We see three reasons for for developing a non-autodifferentiable system.
First, the familiar \texttt{scipy}-based system will serve as an easy entry point for most practitioners who are unfamiliar with autodiff, and may still benefit from and modify the code without expert ML knowledge.
Second, this non-autodiff version serves as a pathfinder and benchmark against which an inevitable autodifferentiable version may be compared.
Finally, the inventory of familiar interpolation algorithms have not yet been ported to PyTorch or JAX, since machine learning or Gaussian Process fitting schemes are generally preferred within the ML community.

@@

    Future work would extend the reach of this study's algorithm to encompass other model grids such as the Sonora series of substellar models, ATLAS \citep{atlas}, and coolTLUSTY \citep{coolTLUSTY}, reaching practitioners studying various types of stars and spectra.
    \texttt{blase} should be able to have an option for the user to input which model grid they would like to base the inference on, and to get even more advanced, perhaps even have the ability to intelligently determine which model grid to use automatically.

@@

    Future work would dive deeper into the science of continuums and develop more advanced methods that can discern continuums with greater accuracy and less modeling restrictions.

@@

\section{Conclusion}
In this study, we have presented a proof-of-concept algorithm that interpolates a subset of the PHOENIX spectral model grid and then uses GP minimization for an inference algorithm.
We create state dictionaries for all spectra in the PHOENIX subset, lossily distilling the spectra with a data compression factor of around 20. These state dictionaries have been uploaded to Zenodo at \url{https://zenodo.org/records/11246174} \citep{zenodo}.
We create the PHOENIX generator and implement a performant spectral reconstruction algorithm, enabling anyone to create reconstructions of PHOENIX spectra with continuously valued stellar parameters.
We introduce and test our inference algorithm, with average absolute deviations from true values of 185 K in $T_\mathrm{eff}$, 0.19 in $\log(g)$, and 0.12 dex in [Fe/H].
In its current state, our algorithm operates on spectra within the PHOENIX subset parameter ranges in \autoref{tab:table1}, requiring that the spectra not contain noticeable Doppler shifting, rotational broadening, or other confounding factors.
The methods discussed here represent a step down a road not traveled in spectral inference, and have the potential to become more advanced in the future by fully utilizing the strengths of physics-informed machine learning.

>>>>

\subsection{Line Identification with \texttt{blase}}
The next step was to convert the PHOENIX subset into a physically interpretable intermediate representation: a table of spectral line properties rather than a\textbf{n} array of fluxes.
We used \texttt{blase}, which models spectral lines as Voigt profiles and tunes the profiles to mimic the original PHOENIX spectrum with back propagation.
Back propagation, put simply, is the process of moving in the autodiff gradient field to update ML model parameters (like mentioned earlier, usually against the gradient because we minimize loss functions).
Four parameters were optimized: the line center $\mu$, the log-amplitude $\ln(a)$, the Gaussian width $\sigma$, and the Lorentzian width $\gamma$.
The optimization used the Adam optimizer \textbf{\citep{adam}} with a learning rate of 0.05 over 100 epochs.
\textbf{\citealt{blase} recommends this setup as the minimum, and for a proof-of-concept implementation, we found it best to leave it as is.}
In addition, we limited two custom parameters: wing cut to 6000 and prominence to 0.005.
Wing cut \textbf{(in pixels)} is a parameter that determines the extent of the Voigt profile to evaluate, saving computational resources by not evaluating negligible line wings.
Prominence \textbf{(in normalized flux counts)} sets a lower limit for the amplitude of detected lines, which saves resources by disregarding shallow lines, so in our case we disregard lines \textbf{with amplitude} $<$ 0.05.
In short, our choices for wing cut and prominence decrease the computational cost of \texttt{blase}'s cloning process at the expense of decreasing its accuracy slightly.
\texttt{blase} uses the pseudo-Voigt approximation, which saves on computational cost while remaining accurate to about 1\% \citep{pseudovoigt}.
The pseudo-Voigt approximation uses a weighted average of a Gaussian and Lorentzian as opposed to a convolution.
\texttt{blase}'s pseudo-Voigt profile implementation uses the following:
\begin{linenomath*}
\begin{subequations}
\begin{gather}
    \tilde{\mathsf{V}}_\mu(\lambda) = a\left[\eta \mathbf{L}(\lambda - \mu'; f) + (1 - \eta)\mathbf{G}(\lambda - \mu'; f)\right]\\
    \eta = \sum_{n=1}^3 \mathbf{u}_n\left(\dfrac{2\gamma}{f}\right)^n\\
    f = 32\sum_{n=0}^5 \mathbf{v}_n \left(\sqrt{2\ln(2)}\sigma\right)^{5-n}(\gamma)^n\\
    \nonumber \mathbf{u} = \mqty[1.36603 \\ -0.47719 \\ 0.11116] \quad \mathbf{v} = \mqty[1 \\ 2.69269 \\ 2.42843 \\ 4.47163 \\ 0.07842 \\ 1]
    \label{eq:equation2}
\end{gather}
\end{subequations}
\end{linenomath*}
where $\mathbf{L}$ and $\mathbf{G}$ are abbreviations for Lorentzian and Gaussian profiles, respectively.
Notice that we use $\mu'$ instead of $\mu$ in the formula.
This is because \texttt{blase} optionally allows the line center to shift slightly during optimization, and it is this shifted center which is used in computation.
The individual Voigt profiles are still indexed by $\mu$ for cross-model line identification, explained in the next section.
Once optimization was complete, the list of identified lines was saved to a `state dictionary': a common representation for pre-trained machine learning models that can be stored to disk for reuse later.
These are stored in the \texttt{.pt} file format for each of the 1,314 PHOENIX subset grid points.
The total disk space these files take up is 465 MB (382 MB when downloaded as a zip archive).
For reference, the storage space the PHOENIX subset takes up on disk is approximately 8.1 GB.
This represents a data compression factor of around 20 just by recasting the spectrum with \texttt{blase}.
\textbf{The state dictionaries are available on Zenodo at \url{https://zenodo.org/records/11246174} \citep{zenodo}.}

@@

Spectral lines were often only detected in some spectra from the PHOENIX subset.
In \autoref{fig:figure2}, we show that different grid points \textbf{show} differing counts of detected spectral lines.
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure2}
    \caption{Number of detected spectral lines at each grid point of a slice of the PHOENIX subset at solar metallicity.
    We can see that the number of detected lines decreases with increasing $T_{\mathrm{eff}}$.
    Also note that from $T_{\mathrm{eff}} = 7000$ K onward, PHOENIX's sampling increment changes from 100 K to 200 K.}
    \label{fig:figure2}
\end{figure*}

Th\textbf{e number of detected spectral lines changes as a function of stellar parameter} for different reasons.
Perhaps for astrophysical reasons: stellar atmospheres genuinely do not produce that line at detectable strength at the given temperature and surface gravity.
Or alternatively, our line-finding and line-association algorithms missed it.
Whatever the cause, these missing lines have immediate practical consequences.
Rectilinear interpolation schemes break in regions where a line does not appear; you can't interpolate a quantity that simply doesn't exist.

To solve this, we artificially populated missing grid points with log-amplitudes of -1000, which retained interpolator stability while nullifying the evaluated line.
Examples of the appearance of missing sections in heatmaps where a line does not appear are shown in \autoref{fig:figure3} and \autoref{fig:figure4}.
In total, across the entire PHOENIX subset, \texttt{blase} detected 128,723 individual spectral lines.
Every one of these lines can be visualized as a manifold mapping a 3D stellar parameter vector to a 4D spectral line parameter vector, and every one \textbf{was} interpolated \textbf{to map stellar parameters to spectral line properties}.
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure3}
    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset slice at solar metallicity of a \textbf{C I} spectral line.
    Notice the missing chunk in the top left of the figure; \texttt{blase} did not detect a spectral line here, but we have to artificially populate those points with lines that have $\ln(a) = -1000$.
    This and all spectral lines shown in this paper were identified using the NIST spectral line database \citep{NIST}.}
    \label{fig:figure3}
\end{figure*}
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure4}
    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset slice at solar metallicity of a He II spectral line.
    We can see that \texttt{blase} detects this line at only a select chunk of grid points in PHOENIX, leading to the large amount of missing data for the line.}
    \label{fig:figure4}
\end{figure*}

\subsection{Continuously Evaluable Manifolds}
For each line, the inputs to the interpolator were the three input parameters $T_{\mathrm{eff}}$, $\log(g)$, and [Fe/H], and the output was a list of four parameters, $\mu'$, $\ln(a)$, $\sigma$, and $\gamma$.
For each line, one of these interpolator objects was created using linear interpolation, and these interpolators were aggregated into a single list, which was then written to disk in the \texttt{.pkl} file storage format.
These interpolators generate multiple manifolds representing the following mapping:
\begin{gather}
    \mqty[T_{\mathrm{eff}} \\ \log(g) \\ \mathrm{[Fe/H]}] \rightarrow \mqty[\mu' \\ \ln(a) \\ \sigma \\ \gamma]
\end{gather}
These interpolators could now be evaluated at any point lying within the domain of the PHOENIX subset, turning a discretely sampled PHOENIX subset into a continuously evaluable function, sometimes called a spectral emulator.
With the given size of the PHOENIX subset, the interpolator list takes up 13.2 GB of disk space.
This evaluation is able to reconstruct an existing PHOENIX spectrum or alternatively interpolate a new spectrum within the domain of the PHOENIX subset, so we call this the PHOENIX generator.
In \autoref{fig:figure5} and \autoref{fig:figure6}, we show the same spectral lines as in \autoref{fig:figure3} and \autoref{fig:figure4}, but now supersampled using the PHOENIX generator evaluated over the same slice.
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure5}
    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ of our \textbf{C I} line vary over the PHOENIX subset slice at solar metallicity, now supersampled with the PHOENIX generator.
    Notice that the missing chunk in the top left still exists and does not display any artifacts, as the artificially populated points are removed after interpolation to retain the model's integrity.
    Also see that the x-axis spacing is now uniform, as the PHOENIX generator was evaluated at constant step.}
    \label{fig:figure5}
\end{figure*}
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure6}
    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ of our He II spectral line vary over the PHOENIX subset slice at solar metallicity, now supersampled with the PHOENIX generator.}
    \label{fig:figure6}
\end{figure*}

@@

\subsection{Spectral Reconstruction Time}
A typical use case for the PHOENIX generator may be to batch reconstruct spectra from an array of input stellar parameters.
Therefore, there is some motivation to reduce the computational cost of this procedure to tractable levels.
We evaluate the computational time needed to use the PHOENIX generator in two distinct ways.
First, for a single input, which would be relevant in serial applications.
Second, for an array of multiple inputs, which would be relevant in parallel applications.
\textbf{\texttt{scipy}'s} \texttt{RegularGridInterpolator} API allows for the passing in of an entire array of input coordinates to be evaluated at once.
However using \texttt{blase} to reconstruct the spectrum from our interpolated state dictionary is always done serially, leading to what is actually more of a pseudo-parallel evaluation, but extremely performant nonetheless.
Performance results are shown in \autoref{fig:figure8}, and we can see that the multi-reconstruction is much faster than a series of single reconstructions.
\textbf{The computer used for this test was a desktop with an AMD Ryzen 9 7950X3D and 32 GB of RAM (spectral reconstruction does not currently leverage the GPU).}

\begin{figure}[t!]
    \centering
    \includegraphics[width=\columnwidth]{figure8}
    \caption{\textbf{Plots} showing the time taken to reconstruct varying numbers of spectra using the PHOENIX generator (lower is better).
    We can see that the time taken per spectr\textbf{um} for the serial implementation hovers around 15 seconds within run-to-run variance, while the pseudo-parallel implementation continually decreases in time taken per spectrum as the number of inputs increases.
    The speedup factor (higher is better) increases as more spectra are generated, which is also a desirable outcome.}
    \label{fig:figure8}
\end{figure}

\section{Bayesian Inference and Testing}
\label{sec:section4}

\subsection{Inference Algorithm}
We elected to use Bayesian optimization as the inference algorithm, specifically the \texttt{gp\_minimize} function from the \texttt{scikit-optimize} library \citep{skopt}.
This algorithm uses a Gaussian Process to model the objective function, which in this case was the RMS (Root-Mean-Square) loss between the interpolated spectrum $\mathsf{M}$ and the true spectrum $\mathsf{D}$, defined as:
\begin{gather}
    \mathcal{L} = \ev{(\mathsf{M} - \mathsf{D})^2}^{1/2}
\end{gather}
The optimizer was configured to first run 100 random evaluations to seed the surrogate model, then run 20 more evaluations now guided by the surrogate model.
This totals to 120 evaluations, a large sample to create a fairly detailed surrogate model, then a moderately precise guided evaluation phase, which was deemed sufficient for this study.
Fine-tuning these numbers is possible, but simply not warranted for a proof-of concept method.
One inference run takes on average just under 7.5 minutes to complete.

@@

We know that in typical observational settings, a coarse estimate for $T_{\mathrm{eff}}$ tends to be fairly well-constrained from ancillary data \textbf{such as photometric color index}.
So when testing the inference algorithm, we limited its search space to only include $T_{\mathrm{eff}}$ which lay within 500 K of the true value on either side.
$\log(g)$ and [Fe/H] were allowed to vary freely.
The test sample consisted of 9 unique $T_{\mathrm{eff}}$ values, 3 unique $\log(g)$ values, and 2 unique [Fe/H] values, totaling 54 unique spectra in the test set.
$T_{\mathrm{eff}}$ ranged from 3000 K to 11000 K in increments of 1000 K, $\log(g)$ ranged from 2 to 6 in increments of 2, and [Fe/H] ranged from -0.5 to 0 in increments of 0.5.

@@

\subsection{Technical Considerations}
To reiterate, the manifold fitting steps are not end-to-end autodifferentiable.
As seen in \autoref{fig:figure1}, these steps rely on \texttt{scipy}, which is not equipped with autodiff.
Without autodifferentiability, the ability to ``machine learn'' is significantly reduced compared to a hypothetical monolithic JAX or PyTorch system.
We see three reasons for for developing a non-autodifferentiable system.
First, the familiar \texttt{scipy}-based system will serve as an easy entry point for most practitioners who are unfamiliar with autodiff, and may still benefit from and modify the code without expert ML knowledge.
Second, this non-autodiff version serves as a\textbf{n initial} benchmark against which an inevitable autodifferentiable version may be compared.
Finally, the inventory of familiar interpolation algorithms have not yet been ported to PyTorch or JAX, since machine learning or Gaussian Process fitting schemes are generally preferred within the ML community.

@@

    Future work would extend the reach of this study's algorithm to encompass other model grids such as the Sonora series of substellar models \textbf{\citep{bobcat, cholla, diamondback, elfowl}}, ATLAS \citep{atlas}, and coolTLUSTY \citep{coolTLUSTY}, reaching practitioners studying various types of stars and spectra.
    \textbf{Future} \texttt{blase} \textbf{versions will} be able to have an option for the user to input which model grid they would like to base the inference on, and to get even more advanced, perhaps even have the ability to intelligently determine which model grid to use automatically.

@@

    Future work would dive deeper into the science of continu\textbf{a} and develop more advanced methods that can discern continu\textbf{a} with greater accuracy and less modeling restrictions.

@@

\section{Conclusion}
In this study, we have presented a proof-of-concept algorithm that interpolates a subset of the PHOENIX spectral model grid and then uses GP minimization \textbf{to infer stellar paremeters}.
We create state dictionaries for all spectra in the PHOENIX subset, lossily distilling the spectra with a data compression factor of around 20. 
The\textbf{y are available on} Zenodo at \url{https://zenodo.org/records/11246174} \citep{zenodo}.
We create the PHOENIX generator and implement a performant spectral reconstruction algorithm, enabling anyone to create reconstructions of PHOENIX spectra with continuously valued stellar parameters.
We introduce and test our inference algorithm, with average absolute deviations from true values of 185 K in $T_\mathrm{eff}$, 0.19 in $\log(g)$, and 0.12 dex in [Fe/H].
In its current state, our algorithm operates on spectra within the PHOENIX subset parameter ranges in \autoref{tab:table1}, requiring that the spectra not contain noticeable Doppler shifting, rotational broadening, or other confounding factors.
The methods discussed here represent a step down a road not traveled in spectral inference, and have the potential to become more advanced in the future by fully utilizing the strengths of physics-informed machine learning.
